On the combination of Set Functions and Failures

### 2013-01-11 docs/src/manual.tex

\lstnewenvironment{curry}{}{}
\newcommand{\listline}{\vrule width0pt depth1.75ex}

@@

Assume you want to implement your own concatenation
for strings in a module \code{String}.
The name and type of this string concatenation should be
\begin{curry}
sconc :: String -> String -> String
\end{curry}
Since the primitive Haskell implementation of this operation
does not now anything about the operational mechanism of Curry
(e.g., needed narrowing, non-deterministic rewriting),
the arguments need to be completely evaluated before
the primitive implementation is called.
This can be easily obtained by the prelude operation \code{(\$\#\#)}
that applies an operation to the \emph{normal form} of the given
argument, i.e., this operation evaluates the argument
to its normal form before applying the operation to it.\footnote{%
There is also a similar prelude operation \code{(\$\#)}
which evaluates the argument only to head-normal form.
This is a bit more efficient and can be used for unstructured
types like \code{Bool}.}
Thus, we define \code{sconc} by
\begin{curry}
sconc :: String -> String -> String
sconc s1 s2 = (prim_sconc $\$$## s1) $\$$## s2

prim_sconc :: String -> String -> String
prim_sconc external
\end{curry}
so that it is ensured that the external operation \code{prim_sconc}
is always called with complete evaluated arguments.

@@

The conversion between Curry types and Haskell types,
i.e., the family of conversion operation \code{fromCurry}
and \code{toCurry}, is defined in the KiCS2 implementation
for all standard data types.
In particular, it is also defined on function types so that one can
easily implement external Curry I/O actions by using
Haskell I/O actions.
For instance, if we want to implement
an external operation to print some string as an output line,
we start by declaring the external operations in the Curry module \code{String}:
\begin{curry}
printString :: String -> IO ()
printString s = prim_printString $\$$## s

prim_printString :: String -> IO ()
prim_printString external
\end{curry}
Next we add the corresponding implementation in the file
\code{External_String.hs} (where \code{C_IO} and \code{OP_Unit} are the names
of the Haskell representation of the Curry type constructor \code{IO}
and the Curry data type \ccode{()}, respectively):
\begin{curry}
external_d_C_prim_printString :: CP.OP_List CP.C_Char -> ConstStore
                              -> CP.C_IO CP.OP_Unit
external_d_C_prim_printString s _ = toCurry putStrLn s
\end{curry}
Here, Haskell's I/O action \code{putStrLn} of type \ccode{String -> IO ()}
is transformed into a Curry I/O action \ccode{toCurry putStrLn}
which has the type
\begin{curry}
CP.OP_List CP.C_Char -> CP.C_IO CP.OP_Unit
\end{curry}
When we compile the Curry module \code{String},
KiCS2 combines these definitions in the target program so that
we can immediately use the externally defined operation
\code{printString} in Curry programs.

>>>>

\lstnewenvironment{curry}{}{}
\lstnewenvironment{currynomath}{\lstset{mathescape=false}}{} % Curry w/o math

\newcommand{\listline}{\vrule width0pt depth1.75ex}

@@

Assume you want to implement your own concatenation
for strings in a module \code{String}.
The name and type of this string concatenation should be
\begin{curry}
sconc :: String -> String -> String
\end{curry}
Since the primitive Haskell implementation of this operation
does not now anything about the operational mechanism of Curry
(e.g., needed narrowing, non-deterministic rewriting),
the arguments need to be completely evaluated before
the primitive implementation is called.
This can be easily obtained by the prelude operation \code{(\$\#\#)}
that applies an operation to the \emph{normal form} of the given
argument, i.e., this operation evaluates the argument
to its normal form before applying the operation to it.\footnote{%
There is also a similar prelude operation \code{(\$\#)}
which evaluates the argument only to head-normal form.
This is a bit more efficient and can be used for unstructured
types like \code{Bool}.}
Thus, we define \code{sconc} by
\begin{currynomath}
sconc :: String -> String -> String
sconc s1 s2 = (prim_sconc $## s1) $## s2

prim_sconc :: String -> String -> String
prim_sconc external
\end{currynomath}
so that it is ensured that the external operation \code{prim_sconc}
is always called with complete evaluated arguments.

@@

The conversion between Curry types and Haskell types,
i.e., the family of conversion operation \code{fromCurry}
and \code{toCurry}, is defined in the KiCS2 implementation
for all standard data types.
In particular, it is also defined on function types so that one can
easily implement external Curry I/O actions by using
Haskell I/O actions.
For instance, if we want to implement
an external operation to print some string as an output line,
we start by declaring the external operations in the Curry module \code{String}:
\begin{currynomath}
printString :: String -> IO ()
printString s = prim_printString $## s

prim_printString :: String -> IO ()
prim_printString external
\end{currynomath} % $
Next we add the corresponding implementation in the file
\code{External_String.hs} (where \code{C_IO} and \code{OP_Unit} are the names
of the Haskell representation of the Curry type constructor \code{IO}
and the Curry data type \ccode{()}, respectively):
\begin{curry}
external_d_C_prim_printString :: CP.OP_List CP.C_Char -> ConstStore
                              -> CP.C_IO CP.OP_Unit
external_d_C_prim_printString s _ = toCurry putStrLn s
\end{curry}
Here, Haskell's I/O action \code{putStrLn} of type \ccode{String -> IO ()}
is transformed into a Curry I/O action \ccode{toCurry putStrLn}
which has the type
\begin{curry}
CP.OP_List CP.C_Char -> CP.C_IO CP.OP_Unit
\end{curry}
When we compile the Curry module \code{String},
KiCS2 combines these definitions in the target program so that
we can immediately use the externally defined operation
\code{printString} in Curry programs.
